---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#",
  fig.width = 10, fig.height = 5,
  fig.align = "center",
  fig.path  = "man/Figure/",
  dev = 'svg'
)
```
# phenofit  
[![Travis Build Status](https://travis-ci.org/kongdd/phenofit.svg?branch=master)](https://travis-ci.org/kongdd/phenofit)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/kongdd/phenofit?branch=master&svg=true)](https://ci.appveyor.com/project/kongdd/phenofit)
[![codecov](https://codecov.io/gh/kongdd/phenofit/branch/master/graph/badge.svg)](https://codecov.io/gh/kongdd/phenofit)
[![License](http://img.shields.io/badge/license-GPL%20%28%3E=%202%29-brightgreen.svg?style=flat)](http://www.gnu.org/licenses/gpl-2.0.html)
[![CRAN](http://www.r-pkg.org/badges/version/phenofit)](https://cran.r-project.org/package=phenofit)
[![total](http://cranlogs.r-pkg.org/badges/grand-total/phenofit)](https://www.rpackages.io/package/phenofit)
[![monthly](http://cranlogs.r-pkg.org/badges/phenofit)](https://www.rpackages.io/package/phenofit)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3605560.svg)](https://doi.org/10.5281/zenodo.3605560)

A state-of-the-art **remote sensing vegetation phenology** extraction package: `phenofit`

 - `phenofit` combine merits of TIMESAT and phenopix
 - A simple and stable growing season dividing methods was proposed
 - Provide a practical snow elimination method, based on Whittaker
 - 7 curve fitting methods and 4 phenology extraction methods
 - We add parameters boundary for every curve fitting methods according to their ecological meaning.
 - `optimx` is used to select best optimization method for different curve fitting methods.


***Task lists***

- [ ] Test the performance of `phenofit` in multiple growing season regions (e.g. the North China Plain);
- [ ] Uncertainty analysis of curve fitting and phenological metrics;
- [x] shiny app has been moved to [phenofit.shiny](https://github.com/kongdd/phenofit.shiny);
- [x] Complete script automatic generating module in shinyapp;
- [x] `Rcpp` improve double logistics optimization efficiency by 60%;
- [x] Support spatial analysis;
- [x] Support annual season in curve fitting;
- [x] flexible fine fitting input ( original time-series or smoothed time-series by rough fitting).
- [x] Asymmetric of Threshold method

![title](man/Figure/Figure1_phenofit_flowchart.svg)   

*<u>Figure 1. The flowchart of phenology extraction in `phenofit`.</u>*

# Installation

You can install phenofit from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("kongdd/phenofit")
```

# Example

Here, we illustrate how to use `phenofit` to extract vegetation phenology from 
MOD13A1 in the sampled points. Regional analysis also can be conducted in the 
similar way.

<!-- ## 1.1 Download MOD13A1 data

Upload point shapefile into GEE, clip MOD13A1 and download vegetation index
data. [Here](https://code.earthengine.google.com/ee3ec39cf3061374dab435c358d008a3) is the corresponding GEE script. 
 -->

## 1.1 Initial weights for input data

Load packages.
```{r load pkg, message=FALSE, echo=FALSE}
library(data.table)
library(magrittr)
library(lubridate)
library(purrr)
library(plyr)
library(ggplot2)
library(phenofit)
```
Set global parameters for `phenofit`
```{r phenofit_parameters}
# lambda   <- 5    # non-parameter Whittaker, only suit for 16-day. Other time-scale
# should assign a lambda.
ymax_min   <- 0.1  # the maximum ymax shoud be greater than `ymax_min` 
rymin_less <- 0.8  # trough < ymin + A*rymin_less
nptperyear <- 23   # How many points for a single year
wFUN       <- wBisquare #wTSM #wBisquare # Weights updating function, could be one of `wTSM`, 'wBisquare', `wChen` and `wSELF`.
```

- Add date according to composite day of the year (DayOfYear), other than image date.
- Add weights according to `SummaryQA`. 

For MOD13A1, Weights can by initialed by `SummaryQA` band (also suit for 
MOD13A2 and MOD13Q1). There is already a `QC` function for `SummaryQA`, i.e. `qc_summary`.

SummaryQA      | Pixel reliability summary QA | weight
---------------| ---------------------------- | ------
-1 Fill/No data| Not processed                | `wmin`
0 Good data    | Use with confidence          | 1
1 Marginal data| Useful but look at detailed QA for more information | 0.5
2 Snow/ice     | Pixel covered with snow/ice  | `wmin`
3 Cloudy       | Pixel is cloudy              | `wmin`

```{r tidy MOD13A1}
data('MOD13A1')
df <- MOD13A1$dt 
st <- MOD13A1$st

df[, `:=`(date = ymd(date), year = year(date), doy = as.integer(yday(date)))]
df[is.na(DayOfYear), DayOfYear := doy] # If DayOfYear is missing
    
# In case of last scene of a year, doy of last scene could in the next year
df[abs(DayOfYear - doy) >= 300, t := as.Date(sprintf("%d-%03d", year+1, DayOfYear), "%Y-%j")] # last scene
df[abs(DayOfYear - doy) <  300, t := as.Date(sprintf("%d-%03d", year  , DayOfYear), "%Y-%j")]

df <- df[!duplicated(df[, .(site, t)]), ]

# MCD12Q1.006 land cover 1-17, IGBP scheme
IGBPnames_006 <- c("ENF", "EBF", "DNF", "DBF", "MF" , "CSH", 
              "OSH", "WSA", "SAV", "GRA", "WET", "CRO", 
              "URB", "CNV", "SNOW", "BSV", "water", "UNC")
# Initial weights
df[, c("QC_flag", "w") := qc_summary(SummaryQA)]
df <- df[, .(site, y = EVI/1e4, t, date, w, QC_flag)]
```

- `add_HeadTail` is used to deal with such situation, e.g. MOD13A2 begins from 2000-02-08. 
We need to construct a series with complete year, which begins from 01-01 for NH, and 07-01 for SH. 
For example, the input data period is 20000218 ~ 20171219. After adding one year in head and 
tail, it becomes 19990101 ~ 20181219. 

## 2.1 load site data
```{r load_data}
sites        <- unique(df$site)
sitename     <- sites[3]
d            <- df[site == sitename] # get the first site data
sp           <- st[site == sitename]

south      <- sp$lat < 0
print      <- FALSE # whether print progress
IsPlot     <- TRUE  # for brks

prefix_fig <- "phenofit"
titlestr   <- with(sp, sprintf('[%03d,%s] %s, lat = %5.2f, lon = %6.2f',
                                     ID, site, IGBPname, lat, lon))
file_pdf   <- sprintf('Figure/%s_[%03d]_%s.pdf', prefix_fig, sp$ID[1], sp$site[1])
```

If need night temperature (Tn) to constrain ungrowing season backgroud value, NA 
values in Tn should be filled.
```{r interp Tn, eval=F}
d$Tn %<>% zoo::na.approx(maxgap = 4)
plot(d$Tn, type = "l"); abline(a = 5, b = 0, col = "red")
```

## 2.2 Check input data
```{r check_input}    
dnew  <- add_HeadTail(d, south, nptperyear = 23) # add additional one year in head and tail
INPUT <- check_input(dnew$t, dnew$y, dnew$w, dnew$QC_flag,
                     nptperyear, south, 
                     maxgap = nptperyear/4, alpha = 0.02, wmin = 0.2)
```

## 2.3 Divide growing seasons

Simply treating calendar year as a complete growing season will induce a considerable error for phenology extraction. A simple growing season dividing method was proposed in `phenofit`.

The growing season dividing method rely on heavily in Whittaker smoother. 

Procedures of initial weight, growing season dividing, curve fitting, and phenology 
extraction are conducted separately.

```{r divide growing season}
par(mar = c(3, 2, 2, 1), mgp = c(3, 0.6, 0))
lambda <- init_lambda(INPUT$y)
# The detailed information of those parameters can be seen in `season`.
# brks   <- season(INPUT, nptperyear,
#                FUN = smooth_wWHIT, wFUN = wFUN, iters = 2,
#                lambda = lambda,
#                IsPlot = IsPlot, plotdat = d,
#                south = d$lat[1] < 0,
#                rymin_less = 0.6, ymax_min = ymax_min,
#                max_MaxPeaksperyear =2.5, max_MinPeaksperyear = 3.5) #, ...
# get growing season breaks in a 3-year moving window
brks2 <- season_mov(INPUT, 
                   FUN = smooth_wWHIT, wFUN = wFUN,
                   maxExtendMonth = 6, r_min = 0.1,
                   IsPlot = IsPlot, IsPlot.OnlyBad = FALSE, print = print)
```

## 2.4 Curve fitting
```{r curve fitting, fig.height=7, fig.align="center"}
fit  <- curvefits(INPUT, brks2,
                  methods = c("AG", "Zhang", "Beck", "Elmore"), #,"klos",, 'Gu'
                  wFUN = wFUN,
                  nextend = 2, maxExtendMonth = 3, minExtendMonth = 1, minPercValid = 0.2,
                  print = print, verbose = FALSE)

## check the curve fitting parameters
l_param <- get_param(fit)
print(str(l_param, 1))
print(l_param$AG)

d_fit <- get_fitting(fit)
## Get GOF information
d_gof <- get_GOF(fit)
# fit$stat <- stat
print(head(d_gof))

# print(fit$fits$AG$`2002_1`$ws)
print(fit$`2002_1`$fFIT$AG$ws)
## visualization
g <- plot_phenofit(d_fit, brks2, NULL, title.ylab = "NDVI", "Time",
                   theme = coord_cartesian(xlim = c(ymd("2000-04-01"), ymd("2017-07-31"))))
grid::grid.newpage(); grid::grid.draw(g)# plot to check the curve fitting
# write_fig(g, "Figure1_phenofit_curve_fitting.pdf", 10, 6)
```

## 2.5 Extract phenology
```{r Extract phenology, fig.height=5, fig.width=8, fig.align="center"}
# pheno: list(p_date, p_doy)
l_pheno <- get_pheno(fit, IsPlot = F) #%>% map(~melt_list(., "meth"))

# ratio = 1.15
# file <- "Figure5_Phenology_Extraction_temp.pdf"
# cairo_pdf(file, 8*ratio, 6*ratio)
# temp <- get_pheno(fit$fits$ELMORE[2:6], IsPlot = T)
# dev.off()
# file.show(file)

## check the extracted phenology
pheno <- get_pheno(fit[1:6], "Elmore", IsPlot = T)
# print(str(pheno, 1))
head(l_pheno$doy$AG)
```

# **References** 


> [1] Kong, D., Zhang, Y., Wang, D., Chen, J., & Gu, X. (2020). Photoperiod Explains the Asynchronization Between Vegetation Carbon Phenology and Vegetation Greenness Phenology. *Journal of Geophysical Research: Biogeosciences*, 125(8), e2020JG005636. https://doi.org/10.1029/2020JG005636
>
> [2] Kong, D., Zhang, Y., Gu, X., & Wang, D. (2019). A robust method for reconstructing global MODIS EVI time series on the Google Earth Engine. *ISPRS Journal of Photogrammetry and Remote Sensing*, 155, 13–24.
>
> [3\] Kong, D., (2020). R package: A state-of-the-art Vegetation Phenology extraction package, `phenofit` version 0.2.6, <https://doi.org/10.5281/zenodo.3605560>
>
> [4\] Zhang, Q., Kong, D., Shi, P., Singh, V.P., Sun, P., 2018. Vegetation phenology on the Qinghai-Tibetan Plateau and its response to climate change (1982–2013). Agric. For. Meteorol. 248, 408–417. <https://doi.org/10.1016/j.agrformet.2017.10.026>

# Acknowledgements

Keep in mind that this repository is released under a GPL2 license, which permits commercial use but requires that the source code (of derivatives) is always open even if hosted as a web service.
