---
title: "phenology based on MOD13 products"
author: "Dongdong Kong"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here, we illustrate how to use `phenofit` in the selected representative points. Regional analysis also can be conducted in the similar way.

## 1. Preparing input data

1.1 Preprocess remote sensing data and download in Google Earth Engine (GEE)  

A 500m buffer was applied to aggregate vegetation index mean value. For a small region (e.g. the 500m buffer), we can consider remote sensing observing condition is same. So, quality control (QC) of point's central grid was used.

1.2 Initial weights for input data. 

For MOD13A1, MOD13A2 and MOD13Q1, Weights can by initialed by `SummaryQA` band.  
Details can be seen in `qc_summary`.


SummaryQA      | Pixel reliability summary QA | weight
---------------| ---------------------------- | ------
-1 Fill/No data| Not processed                | 0
0 Good data    | Use with confidence          | 1
1 Marginal data| Useful but look at detailed QA for more information | 0.5
2 Snow/ice     | Pixel covered with snow/ice  | 0
3 Cloudy       | Pixel is cloudy              | 0

```{r,eval=TRUE}
library(phenofit) 
library(data.table)
library(magrittr)
library(plyr) # for mapvalues

# Global parameters for MOD13A1: 
lambda     <- 5    # The parameter for Whittaker
ymax_min   <- 0.1  # For a growing season, the maximum ymax shoud be greater than `ymax_min` 
nptperyear <- 23   # How many points for a single year
wFUN       <- wTSM # Weights updating function, could be one of `wTSM`, 'wBisquare', `wChen` and `wSELF`. 

# load data
data(dt_MOD13A1) # 16-day time-scale
df <- dt_MOD13A1
df[, w := qc_summary(SummaryQA)]   # Initial weights

# rename variable name as y
varnames <- c("date", "EVI")
I_var    <- match(varnames,colnames(df))
colnames(df)[I_var] <- "y"

# remap SummaryQA factor level, plot_phenofit also use this variable
if ('SummaryQA' %in% colnames(df)){
    df$SummaryQA %<>% factor() %>% mapvalues(c("0", "1", "2", "3"), c(" good", " margin", " snow&ice", " cloud"))
}

sites <- unique(df$site)        # all sites
d     <- df[site == sites[3], ] # get the first site data
# knitr::kable(head(d, 10)) 
```

### 1. Backgroud value
```{r}
d$Tn %<>% na.approx(maxgap = 4)

plot(d$Tn, type = "l")
abline(a = 5, b = 0, col = "red")
```

### 2. Divide growing seasons

Simply treating calendar year as a complete growing season will induce a considerable error for phenology extraction. A simple growing season dividing method was proposed in `phenofit`.

The growing season dividing method rely on heavily in Whittaker smoother. 

Procedures of initial weight, growing season dividing and curve fitting are separated. Phenology extraction and curve fitting are combined together.


```{r}
# 1. Check input data and initial parameters for phenofit
INPUT <- check_input(d$date, d$y, d$w, d$Tn, trim = T, maxgap = nptperyear / 4, alpha = 0.02)

# 2. The detailed information of those parameters can be seen in `season`. 
brks <- season(INPUT, lambda, nptperyear, iters = 3, wFUN = wFUN, IsPlot = F,
                south = d$lat[1] < 0,
                Aymin_less = 0.6, ymax_min = ymax_min, 
                max_MaxPeaksperyear =2.5, max_MinPeaksperyear = 3.5) #, ...

# 3. curve fitting
fit  <- curvefits(INPUT, brks, lambda =lambda, IsPlot = T, 
                     methods = c("AG", "zhang", "beck", "elmore", 'Gu'), #,"klos"
                     nptperyear = nptperyear, debug = F, wFUN = wTSM, 
                     ymax_min = ymax_min, 
                     south = d$lat[1] < 0)
fit$INPUT   <- INPUT
fit$seasons <- brks

plot_phenofit(fit, d) # plot to check the curve fitting

## 3.2 Get GOF information
stat  <- ldply(fit$fits, function(fits_meth){
    ldply(fits_meth, statistic.phenofit, .id = "flag")
}, .id = "meth")

# 4. extract phenology
# pheno: list(p_date, p_doy)
# check extracted phenology extraction for one method.
temp <- ExtractPheno(fit$fits$ELMORE, IsPlot = T)

p <- lapply(fit$fits, ExtractPheno)
pheno  <- map(p, tidyFitPheno, origin = t[1]) %>% purrr::transpose()

fit$stat   <- stat
fit$pheno  <- pheno
```

Calculate all sites in a time.
```{r}
ratio    <- 1.8
lambda   <- 2
sites_sm <- sites[c(1, 3, 4, 6, 9)]
df_sm    <- df[site %in% sites_sm, ]

Cairo::CairoPDF("phenofit_5st_v4.pdf", 10*ratio, 6*ratio)
fits <- dlply(df_sm, .(site), function(d){
    fit <- NA
    tryCatch({
        # Check input data and initial parameters for phenofit
        INPUT <- check_input(d$date, d$y, d$w, trim = T, maxgap = nptperyear / 4, alpha = 0.02)
        # The detailed information of those parameters can be seen in `season`. 
        brks <- season(INPUT, lambda, nptperyear, iters = 3, wFUN = wFUN, IsPlot = F,
                        south = d$lat[1] < 0,
                        Aymin_less = 0.6, ymax_min = ymax_min, 
                        max_MaxPeaksperyear =2.5, max_MinPeaksperyear = 3.5) #, ...
        fit  <- curvefits(INPUT, brks, lambda =lambda, IsPlot = T, 
                          methods = c("AG", "zhang", "beck", "elmore", 'Gu'), #,"klos"
                          nptperyear = nptperyear, debug = F, 
                          wFUN = wFUN, 
                          extend_month = 2, 
                          south = d$lat[1] < 0)
        fit$INPUT   <- INPUT
        fit$seasons <- brks

        titlename <- sprintf('[%s] %s', d$IGBP[1], d$site[1])
        p <- plot_phenofit(fit, d, titlename)
        print(p)
    }, error = function(e){
        message(sprintf("[e] %s:%s", e$message, d$site[1]))
    })
    return(fit)
})
dev.off()
```

You can also call `curvefits2`, which combines all the procedures.   
```{r}
south <- d$lat[1] < 0
fit <- curvefits2(d$date, d$y, d$w, nptperyear = nptperyear,
           wFUN = wFUN, iters = 2,
           lambda = lambda, south = FALSE,
           IsPlot = FALSE,
           Aymin_less = 0.6, ymax_min = ymax_min, 
           methods = c('AG', 'zhang', 'beck', 'elmore', 'Gu'),
           debug = FALSE)
```

Test backgroud value function here

```{r}
profvis::profvis({
        fit <- curvefits2(d$date, d$y, d$w, nptperyear = nptperyear,
                   wFUN = wFUN, iters = 2,
                   lambda = lambda, south = FALSE,
                   IsPlot = FALSE,
                   Aymin_less = 0.6, ymax_min = ymax_min, 
                   methods = c('AG', 'zhang', 'beck', 'elmore', 'Gu')[4],
                   debug = FALSE)
})
```

```{r}
params <- getparam(fit)
print(params)
```

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.
