---
title: "phenology based on MOD13 products"
author: "Dongdong Kong"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here, we illustrate how to use `phenofit` in the selected representative points. Regional analysis also can be conducted in the similar way.

## 1. Preparing input data

1.1 Preprocess remote sensing data and download in Google Earth Engine (GEE)  

A 500m buffer was applied to aggregate vegetation index mean value. For a small region (e.g. the 500m buffer), we can consider remote sensing observing condition is same. So, quality control (QC) of point's central grid was used.

1.2 Initial weights for input data. 

For MOD13A1, MOD13A2 and MOD13Q1, Weights can by initialed by `SummaryQA` band.

SummaryQA      | Pixel reliability summary QA | weight
---------------| ---------------------------- | ------
-1 Fill/No data| Not processed                | 0
0 Good data    | Use with confidence          | 1
1 Marginal data| Useful but look at detailed QA for more information | 0.5
2 Snow/ice     | Pixel covered with snow/ice  | 0
3 Cloudy       | Pixel is cloudy              | 0

```{r}
qc_summary <- function(QA){
    w <- numeric(length(QA)) #default zero
    
    w[QA == 0] <- 1    #clear, good
    w[QA == 1] <- 0.5  #margin
    
    w[QA >= 2 & QA <=3] <- 0 #cloud shadow
    return(w)
}
```

```{r,eval=FALSE}
df_MOD13A1 = fread(paste0(indir, 'flux212_MOD13A1.csv')) #16DAY
MOD13A1[, w := qc_summary(SummaryQA)]
```

### 2. Divide growing seasons

Simply treating calendar year as a complete growing season will induce a considerable error for phenology extraction. A simple growing season dividing method was proposed in `phenofit`.

The growing season dividing method rely on heavily in Whittaker smoother. 


initial parameters for phenofit
```{r}
lambda     <- 5    # The parameter for Whittaker
ymax_min   <- 0.2  # For a growing season, the maximum ymax shoud be greater than `ymax_min` 
nptperyear <- 23   # How many points for a single year
wFUN       <- wTSM # Weights updating function, could be one of `wTSM`, 'wBisquare', `wChen` and `wSELF`. 

# separate every procedure
# t, y, w, nptperyear = 46, lambda, south = FALSE,
INPUT <- check_input(t, y, w, trim = T, maxgap = nptperyear / 4, alpha = 0.02)

# The detailed information of those parameters can be seen in `season`. 
brks  <- season(INPUT, lambda, nptperyear, iters = 3, wFUN = wFUN, IsPlot = F,
                south = south,
                Aymin_less = 0.7,
                max_MaxPeaksperyear =2.5, max_MinPeaksperyear = 3.5, ...) #, ...

fit <- curvefit_site(d$date, d$y, 
                     w = d$w,
                     lambda =lambda,IsPlot = T, 
                     methods = c("AG", "zhang", "beck", "elmore", 'Gu'), #,"klos"
                     nptperyear = nptperyear, debug = F, wFUN = wTSM, 
                     ymax_min = ymax_min, 
                     south = d$lat[1] < 0)
plot_phenofit(fit, d) # plot to check the curve fitting
```

```{r}
params <- getparam(fit)
print(params)
```

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
