---
title: "phenology based on MOD13 products"
author: "Dongdong Kong"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here, we illustrate how to use `phenofit` in the selected representative points. Regional analysis also can be conducted in the similar way.

## 1. Preparing input data

1.1 Preprocess remote sensing data and download in Google Earth Engine (GEE)  

A 500m buffer was applied to aggregate vegetation index mean value. For a small region (e.g. the 500m buffer), we can consider remote sensing observing condition is same. So, quality control (QC) of point's central grid was used.

1.2 Initial weights for input data. 

For MOD13A1, MOD13A2 and MOD13Q1, Weights can by initialed by `SummaryQA` band.

SummaryQA      | Pixel reliability summary QA | weight
---------------| ---------------------------- | ------
-1 Fill/No data| Not processed                | 0
0 Good data    | Use with confidence          | 1
1 Marginal data| Useful but look at detailed QA for more information | 0.5
2 Snow/ice     | Pixel covered with snow/ice  | 0
3 Cloudy       | Pixel is cloudy              | 0

```{r}
qc_summary <- function(QA){
    w <- numeric(length(QA)) # default weight is zero
    
    w[QA == 0] <- 1          # clear, good
    w[QA == 1] <- 0.5        # margin
    
    w[QA >= 2 & QA <=3] <- 0 # Snow/ice, or cloudy
    return(w)
}
```

```{r,eval=TRUE}
library(phenofit) 
library(data.table)
library(magrittr)
library(plyr) # for mapvalues

data(dt_MOD13A1) # 16-day time-scale
df <- dt_MOD13A1
    
df[, w := qc_summary(SummaryQA)]   # Initial weights

# rename variable name as y
varname <- "EVI_500m"
I_var   <- which(colnames(df) == varname[1])
colnames(df)[I_var] <- "y"

# remap SummaryQA factor level, plot_phenofit also use this variable
if ('SummaryQA' %in% colnames(df)){
    df$SummaryQA %<>% factor() %>% mapvalues(c("0", "1", "2", "3"), c(" good", " margin", " snow&ice", " cloud"))
}

sites <- unique(df$site)        # all sites
d     <- df[site == sites[2], ] # get the first site data
knitr::kable(head(d, 10)) 
```

### 2. Divide growing seasons

Simply treating calendar year as a complete growing season will induce a considerable error for phenology extraction. A simple growing season dividing method was proposed in `phenofit`.

The growing season dividing method rely on heavily in Whittaker smoother. 

Procedures of initial weight, growing season dividing and curve fitting are separated. Phenology extraction and curve fitting are combined together.

```{r}
lambda     <- 5    # The parameter for Whittaker
ymax_min   <- 0.2  # For a growing season, the maximum ymax shoud be greater than `ymax_min` 
nptperyear <- 23   # How many points for a single year
wFUN       <- wTSM # Weights updating function, could be one of `wTSM`, 'wBisquare', `wChen` and `wSELF`. 

# Check input data and initial parameters for phenofit
INPUT <- check_input(d$date, d$y, d$w, trim = T, maxgap = nptperyear / 4, alpha = 0.02)

# The detailed information of those parameters can be seen in `season`. 
brks  <- season(INPUT, lambda, nptperyear, iters = 3, wFUN = wFUN, IsPlot = T,
                south = d$lat[1] < 0,
                Aymin_less = 0.6, ymax_min = ymax_min, 
                max_MaxPeaksperyear =2.5, max_MinPeaksperyear = 3.5) #, ...

fit <- curvefits(INPUT, brks, lambda =lambda, IsPlot = T, 
                     methods = c("AG", "zhang", "beck", "elmore", 'Gu'), #,"klos"
                     nptperyear = nptperyear, debug = F, wFUN = wTSM, 
                     ymax_min = ymax_min, 
                     south = d$lat[1] < 0)
plot_phenofit(fit, d) # plot to check the curve fitting
```

```{r}
fit <- curvefits2(d$date, d$y, w = d$w, lambda =lambda,
                             methods = c("AG", "zhang", "beck", "elmore", 'Gu'), #,"klos"
                             nptperyear = nptperyear, debug = F, wFUN = wTSM,
                             ymax_min = ymax_min,
                             south = x$lat[1] < 0)
```


```{r}
params <- getparam(fit)
print(params)
```

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.
