---
title: "Extract vegetation phenology from MOD13A1"
author: "Dongdong Kong"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{phenology_MOD13A1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#",
  fig.width = 10, fig.height = 5,
  fig.align = "center",
  fig.path  = "man/Figure/",
  dev = 'svg'
)
```
## phenofit  
[![Travis Build Status](https://travis-ci.org/kongdd/phenofit.svg?branch=master)](https://travis-ci.org/kongdd/phenofit)
[![codecov](https://codecov.io/gh/kongdd/phenofit/branch/master/graph/badge.svg)](https://codecov.io/gh/kongdd/phenofit)

A state-of-the-art **remote sensing vegetation phenology** extraction package: `phenofit`

 - `phenofit` combine merits of TIMESAT and phenopix
 - A simple and stable growing season dividing methods was proposed
 - Provide a practical snow elimination method, based on Whittaker
 - 7 curve fitting methods and 4 phenology extraction methods
 - We add parameters boundary for every curve fitting methods according to their ecological meaning.
 - `optimx` is used to select best optimization method for different curve fitting methods.

## Installation

You can install phenofit from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("kongdd/phenofit")
```

## Example

Here, we illustrate how to use `phenofit` to extract vegetation phenology from 
MOD13A1 in the sampled points. Regional analysis also can be conducted in the 
similar way.

## 1.1 Download MOD13A1 data

Upload point shapefile into GEE, clip MOD13A1 and download vegetation index
data. [Here](https://code.earthengine.google.com/ee3ec39cf3061374dab435c358d008a3) is the corresponding GEE script. 

## 1.2 Initial weights for input data

Load packages.
```{r load pkg, message=FALSE}
suppressMessages({
    library(data.table)
    library(magrittr)
    library(lubridate)
    library(purrr)
    library(plyr)
    
    library(phenofit)
})
```
Set global parameters for `phenofit`
```{r phenofit_parameters}
# lambda   <- 5    # non-parameter Whittaker, only suit for 16-day. Other time-scale
# should assign a lambda.
ymax_min   <- 0.1  # the maximum ymax shoud be greater than `ymax_min` 
rymin_less <- 0.8  # trough < ymin + A*rymin_less
nptperyear <- 23   # How many points for a single year
wFUN       <- wBisquare #wTSM #wBisquare # Weights updating function, could be one of `wTSM`, 'wBisquare', `wChen` and `wSELF`. 
```

Read the point shapefile to get points coordinate information. Read Enhanced 
Vegetation Index (EVI) exported by `GEE`.

- Add date according to composite day of the year (DayOfYear), other than image date.
- Add weights according to `SummaryQA`. 

For MOD13A1, Weights can by initialed by `SummaryQA` band (also suit for 
MOD13A2 and MOD13Q1). We have written a qc function for `SummaryQA`, `qc_summary`.

SummaryQA      | Pixel reliability summary QA | weight
---------------| ---------------------------- | ------
-1 Fill/No data| Not processed                | `wmin`
0 Good data    | Use with confidence          | 1
1 Marginal data| Useful but look at detailed QA for more information | 0.5
2 Snow/ice     | Pixel covered with snow/ice  | `wmin`
3 Cloudy       | Pixel is cloudy              | `wmin`

```{r tidy MOD13A1}
data('MOD13A1')
df <- MOD13A1$dt 

st <- MOD13A1$st

df[, `:=`(date = ymd(date), year = year(date), doy = as.integer(yday(date)))]
df[is.na(DayOfYear), DayOfYear := doy] # If DayOfYear is missing
    
# In case of last scene of a year, doy of last scene could in the next year
df[abs(DayOfYear - doy) >= 300, t := as.Date(sprintf("%d-%03d", year+1, DayOfYear), "%Y-%j")] # last scene
df[abs(DayOfYear - doy) <  300, t := as.Date(sprintf("%d-%03d", year  , DayOfYear), "%Y-%j")]

df <- df[!duplicated(df[, .(site, t)]), ]

# MCD12Q1.006 land cover 1-17, IGBP scheme
IGBPnames_006 <- c("ENF", "EBF", "DNF", "DBF", "MF" , "CSH", 
              "OSH", "WSA", "SAV", "GRA", "WET", "CRO", 
              "URB", "CNV", "SNOW", "BSV", "water", "UNC")
# Initial weights
df[, c("w", "QC_flag") := qc_summary(SummaryQA)]
# Remap SummaryQA factor level, plot_phenofit use this variable. For other 
# remote sensing data without `SummaryQA`, need to modify `plot_phenofit`

df <- df[, .(site, y = EVI/1e4, t, date, w, QC_flag)]
```

Add one year in head and tail, for growing season dividing. For example, the 
input data period is 20000218 ~ 20171219. After adding one year in head and 
tail, it becomes 19990101 ~ 20181219. 

## 2.1 load data
```{r load_data}
sites        <- unique(df$site)
sitename     <- sites[3]
d            <- df[site == sitename] # get the first site data
sp           <- st[site == sitename]

south      <- sp$lat < 0
print      <- TRUE
IsPlot     <- TRUE # for brks

prefix_fig <- "phenofit"
titlestr   <- with(sp, sprintf('[%03d,%s] %s, lat = %5.2f, lon = %6.2f',
                                     ID, site, IGBPname, lat, lon))
file_pdf   <- sprintf('Figure/%s_[%03d]_%s.pdf', prefix_fig, sp$ID[1], sp$site[1])
```

If need night temperature (Tn) to constrain ungrowing season backgroud value, NA 
values in Tn should be filled.
```{r interp Tn, eval=F}
d$Tn %<>% zoo::na.approx(maxgap = 4)
plot(d$Tn, type = "l"); abline(a = 5, b = 0, col = "red")
```

## 2.1 Check input data
```{r check_input}    
dnew  <- add_HeadTail(d, south, nptperyear = 23) # add additional one year in head and tail
INPUT <- check_input(dnew$t, dnew$y, dnew$w, 
                     nptperyear, south, maxgap = nptperyear/4, alpha = 0.02, wmin = 0.2)
# y0 is used for plot. Original y value has been interpolated and changed.
INPUT$y0 <- dnew$y 
```

## 2.2 Divide growing seasons

Simply treating calendar year as a complete growing season will induce a considerable error for phenology extraction. A simple growing season dividing method was proposed in `phenofit`.

The growing season dividing method rely on heavily in Whittaker smoother. 

Procedures of initial weight, growing season dividing and curve fitting are separated. Phenology extraction and curve fitting are combined together.

```{r divide growing season}
par(mar = c(3, 2, 2, 1), mgp = c(3, 0.6, 0))
lambda <- init_lambda(INPUT$y)
# The detailed information of those parameters can be seen in `season`.
# brks   <- season(INPUT, nptperyear,
#                FUN = wWHIT, wFUN = wFUN, iters = 2,
#                lambda = lambda,
#                IsPlot = IsPlot, plotdat = d,
#                south = d$lat[1] < 0,
#                rymin_less = 0.6, ymax_min = ymax_min,
#                max_MaxPeaksperyear =2.5, max_MinPeaksperyear = 3.5) #, ...
# get growing season breaks in a 3-year moving window
brks2 <- season_3y(INPUT, south = sp$lat[1] < 0, 
                   FUN = wWHIT, wFUN = wFUN,
                   maxExtendMonth = 6, threshold_min = 0.1,
                   IsPlot = IsPlot, print = print, IsOnlyPlotbad = F)
```

## 2.3 Curve fitting
```{r curve fitting, fig.height=7, fig.align="center"}
fit  <- curvefits(INPUT, brks2,
                  methods = c("AG", "zhang", "beck", "elmore"), #,"klos",, 'Gu'
                  debug = F, 
                  wFUN = wFUN,
                  nextent = 2, maxExtendMonth = 3, minExtendMonth = 1,
                  qc = as.numeric(dnew$QC_flag), minPercValid = 0.2,
                  print = print)
fit$INPUT   <- INPUT
fit$seasons <- brks2

## check the curve fitting parameters
params <- getparam(fit)
print(str(params, 1))
print(params$AG)

## Get GOF information
stat  <- ldply(fit$fits, function(fits_meth){
    ldply(fits_meth, statistic.phenofit, .id = "flag")
}, .id = "meth")
fit$stat <- stat
print(head(stat))

print(fit$fits$AG$`2002_1`$ws)
## visualization
# svg("Figure1_phenofit_curve_fitting.svg", 11, 7)
# Cairo::CairoPDF(file_pdf, 11, 6) #
# dev.off()
g <- plot_phenofit(fit, d, titlestr)
grid::grid.newpage(); grid::grid.draw(g)# plot to check the curve fitting
```

## 2.4 Extract phenology.
```{r Extract phenology, fig.height=5, fig.width=8, fig.align="center"}
# pheno: list(p_date, p_doy)
p <- lapply(fit$fits, ExtractPheno)
pheno  <- map(p, tidyFitPheno, origin = INPUT$t[1]) %>% purrr::transpose()
fit$pheno  <- pheno

# ratio = 1.15
# file <- "Figure5_Phenology_Extraction_temp.pdf"
# cairo_pdf(file, 8*ratio, 6*ratio)
# temp <- ExtractPheno(fit$fits$ELMORE[2:6], IsPlot = T)
# dev.off()
# file.show(file)

## check the extracted phenology
temp <- ExtractPheno(fit$fits$ELMORE[1:6], IsPlot = T, TRS = 0.5)

print(str(pheno, 2))
head(pheno$doy$AG)
```
